A Non-local Bayesian Video Denoising Method

### 2015-12-10 doc/201507_tip/paper.tex
[doc] TIP paper: second draft sent to JMM with description of negative weights

% pablo's commands for comments and suggested modifications
\newcommand{\pa}[1]{\textcolor{cyan}{#1}}
\newcommand{\psub}[2]{\pa{\sout{#1}}\pa{/#2}}
\newcommand{\prm}[1]{\pa{\sout{#1}}}
\newcommand{\padd}[1]{\pa{#1}}
\newcommand{\pcomment}[1]{\textit{\pa{#1}}}

@@

A natural next step is to consider 3D spatio-temporal patches. This has two
main advantages (at the expense of a higher computational cost). First it
provides a mechanism to impose some temporal consistency on the result, since
now, the temporally continguous 2D spatial slices of a spatio-temporal patch
are filtered jointly. Secondly, a 3D patch has a higher number of pixels
resulting in a reduction of the noise in the distance between patches. On the
other hand, to get the same reduction in the distance noise with 2D patches one
needs to increase their spatial size which leads to a smaller number of similar
patches.

@@

Motion-compensated patches have the benefit that their spatial slices are very
similar between each other (or the same, assuming purely translational pixelic
motion). As a consequence, temporal filtering can be easily performed by
filtering inside the patch in the temporal direction. 
Because of this, methods that use motion-compensated 3D patches only look for
similar motion-compensated 3D patches starting in the same frame.

%It has been shown by \cite{Liu2010} and \cite{Maggioni2011,Maggioni2012} that motion estimation can still be
%beneficial for patch-based methods. 
%
An example of this is the method presented in \cite{Liu2010}. The authors
propose a method based on non-local means. The method is described in terms of
2D patches. To denoise a target 2D patch, the $k$ nearest neighbors in the current
frame are found. The 2D patches in their forward and backward trajectories are
used also to denoising the target patch. As a consequence the set of similar
patches averaged to denoise each pixel varies smoothly along motion
trajectories. This yields results with high temporal consistency. 
To estimate motion, their method resorts to a robust optical flow algorithm
based on \cite{Bruhn2005}.

@@

In this work we present an extension to video of the non-local Bayes image
denoising method \cite{Lebrun2013a,Lebrun2013ipol} using 3D rectangular
patches and which does not require motion estimation. Yet, it provides
results which compare favourably to the current state-of-the-art,
both quantitatively by a significant marging and qualitatively, specifically in
terms of the temporal consistency attained.
The method works by learning a Gaussian \emph{a priori} model from sets of 
similar noisy patches. Each patch is denoised by computing the \emph{maximum a
posteriori} (MAP) with the estimated prior.
Learning the Gaussian prior amounts to estimating a covariance matrix with a
limited number of samples in a high dimensional space. Motivated by recent 
theoretical results and supported by an empirical analysis, we add the 
assumtion that the \emph{a priori} model is of low rank (\ie the clean patches
lie in a lower dimensional subspace). This assumtion lowers the computational cost
of the algorithm, and improves the results.

@@

The rest of the paper is organized as follows: in \S
\ref{sec:review_nonlocal_Bayes} we review the Bayesian patch denoising principle 
and describe our extension to video. The final algorithm and details about its
implementation are given in \S \ref{sec:algorithm}.
Some relevant aspects of the proposed method are discussed in \S
\ref{sec:discussion}.
In \S \ref{sec:parameters}
we explain the selection of the parameters. Some results, including a
comparison with the state-of-the-art methods, are shown in \S
\ref{sec:results}. Concluding remarks are given in \S \ref{sec:conclusion}.

@@

A patch based Bayesian algorithm can assume that the patches similar to a given
patch follow a Gaussian distribution.
The problem 
of denoising can then be formulated as a problem of optimal Bayesian inference, where the
parameters of the prior are learnt from the noisy data.

@@

Given a rectangular 3D patch $\ma q$ of size $s_x\times s_x\times s_t$ of the noisy video $v$, and the 
corresponding patch $\ma p$ of the clean video $u$, we assume the following 
Gaussian linear model of $\ma p$ and $\ma q$:
\begin{align}
	\mathds{P}(\ma p) &= \mathcal N(\overline {\ma p}, C) \propto \exp\left(-\frac12\langle \ma p - \overline{\ma p}, C^{-1}(\ma p - \overline{\ma p})\rangle\right) \label{eq:prior}\\
	\mathds{P}(\ma q|\ma p) &= \mathcal N(\ma p, \sigma I) \propto \exp\left(-\frac1{2\sigma^2}\|\ma q - \ma p\|^2\right) \label{eq:obs}
\end{align}
Once the mean patch $\overline{\ma p}$ and the
covariance matrix $C$ have been estimated from the image itself, the MAP
estimate $\widetilde{\ma p}$ given a noisy patch $\ma q$ is obtained as:
\[ \widetilde{\ma p} = \argmax_{\ma p} \mathds P(\ma p | \ma q) = \argmin_{\ma p} -\log \mathds P(\ma p | \ma q). \]
Since we are working with Gaussian distributions, the log-posterior is a quadratic (positive definite)
function and the MAP can be computed explicitly as (see  \cite{Lebrun2013a})
\begin{equation}
	\widetilde{\ma p} = \overline{\ma p} + C(C + \sigma^2 I)^{-1}(\ma q - \overline{\ma p})
	\label{eq:map}
\end{equation}
The matrix in the above equation can be diagonalized in the basis of
principal directions of the covariance matrix $C$.
Let $C = U\Lambda U^T$ denote the eigendecomposition of $C$. Then we have that
\[U^T(\widetilde{\ma p} - \overline{\ma p}) \,\,\,\, = \,\,\,\, 
	\Lambda(\Lambda + \sigma^2 I)^{-1}\,\,\,\,
	U^T (\ma q - \overline{\ma p}).\]
The diagonal operator $W = \Lambda(\Lambda + \sigma^2I)^{-1}$
has the following coefficients:
\begin{equation}
w_{ii} = \frac{\lambda_i}{\lambda_i + \sigma^2}
	%	       = (1 + \sigma^2/\lambda_i)^{-1}
				 = (1 + \text{snr}_i^{-1})^{-1}
				 \label{eq:wiener_coeffs}
\end{equation}
which correspond to a Wiener filter on the principal components.

@@

The parameters of the \textit{a priori} model, are learnt from the noisy video as
follows. For each noisy patch $\ma q$, we select the most $n$
similar 3D patches found in a search region centered at $\ma q$. The
search region is a spatio-temporal rectangle of size $w_x \times
w_x \times w_t$ ($w_x, w_t$ are odd numbers).

@@

We will also assume that the covariance matrix has a lower rank $k$.  In this
case, it has been shown REF PPCA that maximum likelihood estimate is given by
truncated the spectral decomposition of $\widehat C$ to the $k$ largest
eigenvalues.
\begin{equation}
\widehat C_k = \sum_{i = 1}^k\widehat \lambda_i\widehat U_i\widehat U_i^T,
\end{equation}
where $\widehat \lambda_i$, $\widehat U_i$ denote the $i$-th largest
eigenvalue and its corresponding eigenvector respectively.

The Figures \ref{fig:patch_groups_pos} and \ref{fig:patch_groups_patches}
show two examples of groups of similar $9\times 9\times 4$ patches taken from
a sequence contaminated with AGWN of $\sigma = 40$. Figure
\ref{fig:patch_groups_pos} shows the position of 200 nearest neighbors of a
reference patch searched in a $37\times 37\times 5$ search region. The
reference patch is centered in the central frame. The position of the first 5
similar patches are highlighted in red, and they follow the trajectory of the
reference patch throughout the search region.  The next 40 similar patches are
shown in red. In the first example they are spatially shifted versions of the
five nearest neighbors. In the second example, in addition to these shifts
other similar structures in the red sign are detected.  The next nearest
neighbors are shown in blue.

Figure \ref{fig:patch_groups_patches} shows the 20 nearest neighbors in 
more detail. Each patch is displayed by vertically stacking its four $9\times
9$ spatial slices. For the two examples shown in Figure \ref{fig:patch_groups_pos}
the figure shows the noisy patches, the result of the MAP estimation (setting
$r = 40$) and the mean patch together with the leading 19 principal directions.
Since the patch is not motion-compensated, all spatial slices can be different,
and thus the same applies to the principal directions.
The spatial slices of the principal directions show oscillating patterns, with
increasing frequency. In some cases these patterns are localized in a part of
the spatial slice.

@@

Other methods for image denoising have been proposed in the
literature based on Gaussian models for the patch density. In \cite{Yu2012},
the authors introduce a framework for solving inverse problems, based on the
assumption that the patches of an image are distributed according to a Gaussian
Mixture Model (GMM). A method is proposed to learn the parameters of the model from
the image. A GMM is also used in \cite{Zoran2011}, but it is trained from a
database of $2\cdot 10^6$ patches randomly sampled from the Berkeley database.

@@

\begin{figure*}[htpb!]
	\centering
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}\\
	\vspace{.2cm}
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
%	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
	\caption{Examples of sets of 200 nearest neighbors spatio-temporal patches,
	for patches of size $9\times 9\times 4$.
	The images in each row show the five frames of a $37\times37\times5$ search region. 
	In red we show the positions of the 5 nearest neighbors, in green the next 40 and in blue the rest.
	Note the points shown correspond to the top-left pixel of the first spatial
	slice of each patch.
	The reference patch is the red patch in the center of the search region. To highlight the
	position of the patches, the color of the images has been attenuated.}
	\label{fig:patch_groups_pos}
\end{figure*}

\begin{figure}[htpb!]
	\centering
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
%	\vspace{.1cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
	\vspace{.2cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
%	\vspace{.1cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
	\vspace{.2cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
%	\vspace{.1cm}
	\caption{For each of the two sets of patches of Fig. \ref{fig:patch_groups_pos} we show
	the 20 nearest neighbors (top), the corresponding MAP estimates (middle) and the sample mean 
	and first 19 principal directions computed from the sample covariance matrix (bottom).
	For each image, each column corresponds to the 4 spatial slices of a $9\times9\times4$ patch.}
	\label{fig:patch_groups_patches}
\end{figure}

@@

\paragraph{Handling of color} Color videos are handled as in \cite{Lebrun2013a}.
%
In the first stage, we express the video in a luminance-chrominance colorspace,
specifically we apply the opponent color transform (see \cite{Dabov2007tip,Lebrun2013a}).
Patch distances are computed using the luminance only. Using the 
$n$ most similar patches, a group is built for each channel. These
groups are filtered independently (a Gaussian model is learnt for each of them).
%
In the second stage, the patch distance is computed using the RGB patches, and
a single \textit{a priori} Gaussian model is built for the each set
of similar RGB patches.

@@

\paragraph{Patch distance threshold} When building the group of similar
patches during the second stage, more than $n$ patches are allowed
if their distances with respect to the reference patch are smaller than a
threshold $\tau$.  As in \cite{Lebrun2013ipol}, we set $\tau = 4$ (the distance between
patches is normalized by the number of elements in the patch and the number of
channels).

@@

Let us consider the complexity of constructing and filtering a group of
$n$ similar patches.
Processing each group of similar patches is of asymptotic order $\mathcal
O(w_x^2 w_t s_x^2 s_t + ns_x^4s_t^2 + s_x^4s_t^2r)$.
The first term comes from the exhaustive search of the $n$ nearest patches in
the search window.  The remaining terms correspond respectively to the
computation of the covariance matrix and its truncated eigen-decomposition.

The most expensive operations are the computation of the covariance matrix and
of its principal directions.
Let us note that there are methods for computing the $r$ principal directions
of a set of points that do not require computing the covariance matrix
\cite{Halko2011}. Instead these methods efficiently approximate a truncated SVD
of the data matrix, and would 
allow to improve the computational complexity to $\mathcal O(w_x^2 w_t s_x^2
s_t + s_x^2s_trn)$.

The computation can be substantially accelerated by reducing the 
number of groups of patches that are jointly processed.
Each group of patches is associated to the reference patch from which 
the nearest neighbors are computed.
We apply two tricks
considered in \cite{Lebrun2013ipol} in order to reduce the number of reference patches. 
The first one consists on visiting only 
those patches on a coarser spatial grid with vertical and horizontal separation of $s_x/2$
(similar to \cite{Dabov2007tip}). This reduces the number of reference patches by a
factor of $s_x^2/4$. The second trick consists on the following.  After
filtering the patches in a group, these patches are not considered again
as reference patches. 
Note that we ensure that each reference patch is filtered at least once. 
Due to the spacing of the coarse grid, this guarantees that each pixel is
overlapped by four reference patches on the same frame, so it will receive at
least four contributions in the aggregation. The actual number of aggregated
estimates for each pixel is much higher, considering the reference and
non-reference patches (which can also be filtered) in the neighboring frames.

@@

Let us assume that 
the brightness of the original video is constant along motion trajectories, so
that $u(x,t + \tau) = u(\varphi(x,t + \tau; -\tau),t)$, where
$\varphi(x,t+\tau;\, \cdot\, )$ denotes the motion trajectory passing though
$(x,t + \tau)$.
Furthermore, let us assume that the motion in the video is approximately
translational.
When the motion is known or can be estimated, then it is natural to consider
motion compensated spatio-temporal patches. 
With the assumptions we are considering, all spatial slices of a $s_x\times s_x\times s_t$ motion-compensated patch
coincide, so the distance between two spatio-temporal patches is proportional to
the distance of a 2D spatial patch of size $s_x\times s_x$.
%
As a consequence, by increasing $s_t$, on can reduce the 
noise in the patch distance and have the same pattern of self-similarity as a 
2D patch.

@@

Learning the parameters of the \emph{a priori} distribution for each group of
noisy patches amounts to estimating a covariance matrix is a high dimensional space 
and with a limited number of samples. 
%The latter is controlled by the $n$ parameter, which
%should not be too high in order for the Gaussian prior to be valid.
%
The empirical covariance matrix $\widehat C$ converges to the true covariance matrix
$C$ as the number of samples from the underlying Gaussian distribution grows to infinity.
But in our case, the number of samples we can ``draw'' from the Gaussian is limited:
%The Gaussian is only an approximate density model for the set of
%$n$ nearest neighbors of a target patch.
As we increase the number $n$ of neighbors of a patch, we include patches
that are not as similar to the reference patch and the Gaussian model becomes less valid.

Thus there is a trade-off between accuracy of Gaussian model and precision of the
covariance matrix. We will find a compromise to this trade-off empirically  in
\S \ref{sec:parameters}. But it still intersting to review the literature 
on the sample covariance matrix as an estimator of the population covariance.

>>>>

% pablo's commands for comments and suggested modifications
\newcommand{\pa}[1]{\textcolor{cyan}{#1}}
\newcommand{\psub}[2]{\pa{\sout{#1}}\pa{/#2}}
\newcommand{\prm}[1]{\pa{\sout{#1}}}
\newcommand{\padd}[1]{\pa{#1}}
%\newcommand{\pcomment}[1]{\textit{\pa{#1}}}
\newcommand{\pcomment}[1]{}

@@

An interesting alternative was proposed in \cite{Liu2010},
based on non-local means.
To denoise a target 2D patch in a frame, its forward and backward trajectories are
determined using a robust optical flow algorithm based on \cite{Bruhn2005}. 
Each half of the trajectory extends for $t$ frames. There are $2t + 1$ 2D patches
centered at each point of the trajectory.
For each of these patches, the $n$ nearest neighbors are found
using an efficient approximate searching strategy. The target 2D patch is denoised
by performing a weighted average of these $(2t + 1)n$ patches. The weight considers
the patch similarity as in the original non-local means method, multiplied by a coefficient
that decreases smoothly with the temporal distance between a source patch and the target patch.
%
In this way, 2D patches in the same trajectory are denoised using a smoothly
varying non-local average of similar 2D patches. The result is a video with
a high temporal consistency.

A natural next step is to consider 3D spatio-temporal patches. This has two
main advantages (at the expense of a higher computational cost). First it
provides a mechanism to impose some temporal consistency on the result, since
now, the temporally contiguous 2D spatial slices of a spatio-temporal patch
are filtered jointly. Secondly, a 3D patch has a higher number of pixels
resulting in a reduction of the noise in the distance between patches. On the
other hand, to get the same reduction in the distance noise with 2D patches one
needs to increase their spatial size which leads to a smaller number of similar
patches.

@@

Motion-compensated patches have the benefit that their spatial slices are very
similar between each other. As a consequence, temporal filtering can be easily performed by
filtering inside the patch in the temporal direction. 
Because of this, methods that use motion-compensated 3D patches only look for
similar motion-compensated 3D patches starting in the same frame.

@@

In this work we present new Bayesian patch-based video denoising method using
3D rectangular patches and not requiring motion estimation.
Yet, it provides
results which compare favourably to the current state-of-the-art,
both quantitatively by a significant margin and qualitatively, specifically in
terms of the temporal consistency attained.
In the spirit of  \cite{Lebrun2013a,Lebrun2013ipol},
the method works by learning a Gaussian \emph{a priori} model from sets of 
similar noisy patches. Each patch is denoised by computing the \emph{maximum a
posteriori} (MAP) with the estimated prior.
Learning the Gaussian prior amounts to estimating a covariance matrix with a
limited number of samples in a high dimensional space. Motivated by recent 
theoretical results and supported by an empirical analysis, we add the 
assumption that the \emph{a priori} model is of low rank (\ie the clean patches
lie in a lower dimensional subspace). This assumption lowers the computational cost
of the algorithm, and improves the results.

@@

The rest of the paper is organized as follows: in \S
\ref{sec:review_nonlocal_Bayes} we review our patch-based Bayesian optimal estimator for
video denoising. The final algorithm and details about its implementation are
given in \S \ref{sec:algorithm}.
Some relevant aspects of the proposed method are discussed in \S
\ref{sec:discussion}.
In \S \ref{sec:parameters}
we explain the selection of the parameters. Some results, including a
comparison with the state-of-the-art methods, are shown in \S
\ref{sec:results}. Concluding remarks are given in \S \ref{sec:conclusion}.

@@

Our fundamental assumption is that video patches similar to a given
patch follow a Gaussian distribution.
The problem 
of denoising can then be formulated as a problem of optimal Bayesian inference, where the
parameters of the prior are learnt from the noisy data.

@@

Given a rectangular 3D patch $\ma q$ of size $s_x\times s_x\times s_t$ of the noisy video $v$, and the 
corresponding patch $\ma p$ of the clean video $u$, we assume the following 
Gaussian linear model of $\ma p$ and $\ma q$:
\begin{align}
	\mathds{P}(\ma p) &= \mathcal N(\overline {\ma p}, C) \propto \exp\left(-\frac12\langle \ma p - \overline{\ma p}, C^{-1}(\ma p - \overline{\ma p})\rangle\right) \label{eq:prior}\\
	\mathds{P}(\ma q|\ma p) &= \mathcal N(\ma p, \sigma I) \propto \exp\left(-\frac1{2\sigma^2}\|\ma q - \ma p\|^2\right) \label{eq:obs}
\end{align}
Once the mean patch $\overline{\ma p}$ and the
covariance matrix $C$ have been estimated from the image itself, the MAP
estimate $\widetilde{\ma p}$ given a noisy patch $\ma q$ is obtained as:
\[ \widetilde{\ma p} = \argmax_{\ma p} \mathds P(\ma p | \ma q) = \argmin_{\ma p} -\log \mathds P(\ma p | \ma q). \]
Since we are working with Gaussian distributions, the log-posterior is a quadratic (positive definite)
function and the MAP can be computed explicitly as (see  \cite{Lebrun2013a})
\begin{equation}
	\widetilde{\ma p} = \overline{\ma p} + C(C + \sigma^2 I)^{-1}(\ma q - \overline{\ma p})
	\label{eq:map}
\end{equation}
The matrix in the above equation can be diagonalized in the basis of
principal directions of the covariance matrix $C$.
Let $C = U\Lambda U^T$ denote the spectral decomposition of $C$. Then we have that
\[U^T(\widetilde{\ma p} - \overline{\ma p}) \,\,\,\, = \,\,\,\, 
	\Lambda(\Lambda + \sigma^2 I)^{-1}\,\,\,\,
	U^T (\ma q - \overline{\ma p}).\]
The diagonal operator $W = \Lambda(\Lambda + \sigma^2I)^{-1}$
has the following coefficients:
\begin{equation}
w_{ii} = \frac{\lambda_i}{\lambda_i + \sigma^2}
	%	       = (1 + \sigma^2/\lambda_i)^{-1}
				 = (1 + \text{snr}_i^{-1})^{-1}
				 \label{eq:wiener_coeffs}
\end{equation}
which correspond to a Wiener filter on the principal components.

@@

The parameters of the \textit{a priori} model are learnt from the noisy video as
follows. For each noisy patch $\ma q$, we select the most $n$
similar 3D patches found in a search region centered at $\ma q$. The
search region is a spatio-temporal rectangle of size $w_x \times
w_x \times w_t$ ($w_x, w_t$ are odd numbers).

@@

We will also assume that the covariance matrix has a lower rank $k$.  In this
case, it has been shown \cite{Tipping1999} that the maximum likelihood estimate is given by
truncating the spectral decomposition of $\widehat C$ to the $k$ largest
eigenvalues 
\begin{equation}
\widehat C_k = \sum_{i = 1}^k\widehat \lambda_i\widehat U_i\widehat U_i^T,
\end{equation}
where $\widehat \lambda_i$, $\widehat U_i$ denote the $i$-th largest
eigenvalue and its corresponding eigenvector respectively.

\padd{The estimated eigenvalues are given by $\widehat \lambda_i = \widehat
\nu_i - \sigma^2$, where $\widehat \nu_i$ are the eigenvalues of the empirical
covariance matrix of the noisy patches. In practice, some of the smallest
estimated eigenvalues $\widehat \nu_i$ can be smaller than $\sigma^2,$ causing
$\widehat \lambda_i$ and the corresponding filter coefficient $w_{ii}$ to be
negative. Negative variances do not have physical meaning, and the
corresponding weights have a negative impact on the MAP estimate. This is an
issue of empirical Wiener filters and it is typically addressed by zeroing out
the negative coefficients. Later in \S
\ref{sec:parameters} we compare the results obtained by zeroing out negative
coefficients against using them.}

Figures \ref{fig:patch_groups_pos} and \ref{fig:patch_groups_patches}
show two examples of groups of similar $9\times 9\times 4$ patches taken from
a sequence contaminated with AGWN of $\sigma = 40$. Figure
\ref{fig:patch_groups_pos} shows the position of 200 nearest neighbors of a
reference patch searched in a $37\times 37\times 5$ search region. The
reference patch is centered in the central frame. The position of the first 5
similar patches are highlighted in red, and they follow the trajectory of the
reference patch throughout the search region.  The next 40 similar patches are
shown in red. In the first example they are spatially shifted versions of the
five nearest neighbors. In the second example, in addition to these shifts
other similar structures in the red sign are detected.  The next nearest
neighbors are shown in blue.

Figure \ref{fig:patch_groups_patches} shows the 20 nearest neighbors in 
more detail. Each patch is displayed by vertically stacking its four $9\times
9$ spatial slices. For the two examples shown in Figure \ref{fig:patch_groups_pos}
the figure shows the noisy patches, the result of the MAP estimation and the
mean patch together with the leading 19 principal directions. The Gaussian
models were learnt from the $n = 200$ similar patches shown in Figure
\ref{fig:patch_groups_pos}. The MAP estimates were computed setting $r = 40$.
Since the patch is not
motion-compensated, all spatial slices can be different,
and thus the same applies to the principal directions.
The spatial slices of the principal directions show oscillating patterns, with
increasing frequency. In some cases these patterns are localized in a part of
the spatial slice.

@@

Other methods for image denoising have been proposed in the
literature based on Gaussian models for the patch density. In \cite{Yu2012},
the authors introduce a framework for solving inverse problems, based on the
assumption that the patches of an image are distributed according to a Gaussian
Mixture Model (GMM) which is learnt from
the image. A GMM is also used in \cite{Zoran2011}, but it is trained from a
database of $2\cdot 10^6$ patches randomly sampled from the Berkeley database.

@@

\begin{figure*}[htpb!]
	\centering
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}\\
	\vspace{.2cm}
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
%	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
	\caption{Examples of sets of 200 nearest neighbors spatio-temporal patches,
	for patches of size $9\times 9\times 4$.
	The images in each row show the five frames of a $37\times37\times5$ search region. 
	In red we show the positions of the 5 nearest neighbors, in green the next 40 and in blue the rest.
	Note that the points shown correspond to the top-left pixel of the first spatial
	slice of each patch.
	The reference patch is the red patch in the center of the search region. To highlight the
	position of the patches, the color of the images has been attenuated.}
	\label{fig:patch_groups_pos}
\end{figure*}

\begin{figure}[htpb!]
	\centering
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
%	\vspace{.1cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
	\vspace{.2cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
%	\vspace{.1cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
	\vspace{.2cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
%	\vspace{.1cm}
	\caption{For each of the two sets of patches of Fig. \ref{fig:patch_groups_pos} we show
	the 20 nearest neighbors (top), the corresponding MAP estimates (middle) and the sample mean 
	and first 19 principal directions computed from the sample covariance matrix (bottom).
	For each image, each column corresponds to the 4 spatial slices of a $9\times9\times4$ patch.}
	\label{fig:patch_groups_patches}
\end{figure}

@@

\paragraph{Handling of color} %Color videos are handled as in \cite{Lebrun2013a}.
%
In the first stage, we express the video in a luminance-chrominance color space,
specifically we apply the opponent color transform (see \cite{Dabov2007tip,Lebrun2013a}).
Patch distances are computed using the luminance only. Using the 
$n$ most similar patches, a group is built for each channel. These
groups are filtered independently (a Gaussian model is learnt for each of them).
%
In the second stage, the patch distance is computed using the RGB patches, and
a joint \textit{a priori} Gaussian model is built for each set
of similar RGB patches. Note that in first stage, we learn three Gaussian models
of dimensionality $s_x^2s_t$, whereas in the second stage we learn a single 
Gaussian model of tripled dimensionality.

@@

\paragraph{Patch distance threshold} When building the group of similar
patches during the second stage, more than $n$ patches are allowed
if their distances with respect to the reference patch are smaller than a
threshold $\tau$.  As in \cite{Lebrun2013ipol}, we set $\tau = 4$ (the distance between
patches is normalized by the number of elements in the patch and the number of
channels and the values of each channel are in the range  $[0,255]$).

@@

Let us consider the complexity of constructing and filtering a group of
$n$ similar patches.
Processing each group of similar patches is of asymptotic order $\mathcal
O(w_x^2 w_t s_x^2 s_t + ns_x^4s_t^2 + s_x^4s_t^2r)$.
The first term comes from the exhaustive search of the $n$ nearest patches in
the search window.  The remaining terms correspond respectively to the
computation of the covariance matrix and its truncated spectral decomposition.

The most expensive operations are the computation of the covariance matrix and
of its principal directions.
There are methods for computing the $r$ principal directions
of a set of points that do not require computing the covariance matrix
\cite{Halko2011}. Instead these methods efficiently approximate a truncated SVD
of the data matrix, and would 
allow to improve the computational complexity to $\mathcal O(w_x^2 w_t s_x^2
s_t + s_x^2s_trn)$.

The computation can be substantially accelerated by reducing the 
number of groups of patches that are jointly processed.
Each group of patches is associated to the reference patch from which 
the nearest neighbors are computed.
We apply two tricks
considered in \cite{Dabov2007tip,Lebrun2013ipol} in order to reduce the number of reference patches. 
The first one consists on visiting only 
those patches on a coarser spatial grid with vertical and horizontal separation
of $s_x/2$. This reduces the number of reference patches by a factor of
$s_x^2/4$. The second trick consists on the following.  After
filtering the patches in a group, these patches are not considered again
as reference patches. 
Note that we ensure that each reference patch is filtered at least once. 
Due to the spacing of the coarse grid, this guarantees that each pixel is
overlapped by four reference patches on the same frame, so it will receive at
least four contributions in the aggregation. The actual number of aggregated
estimates for each pixel is much higher, considering the reference and
non-reference patches (which can also be filtered) in the neighboring frames.

@@

Let us assume that 
the brightness of the original video is constant along motion trajectories, so
that $u(x,t + \tau) = u(\varphi(x,t + \tau; -\tau),t)$, where
$\varphi(x,t+\tau;\, \cdot\, )$ denotes the motion trajectory passing through
$(x,t + \tau)$.
Furthermore, let us assume that the motion in the video is approximately
translational.
When the motion is known or can be estimated, then it is natural to consider
motion compensated spatio-temporal patches. 
With the assumptions we are considering, all spatial slices of a $s_x\times s_x\times s_t$ motion-compensated patch
coincide, so the distance between two spatio-temporal patches is proportional to
the distance of a 2D spatial patch of size $s_x\times s_x$.
%
As a consequence, by increasing $s_t$, one can reduce the 
noise in the patch distance and have the same pattern of self-similarity as a 
2D patch.

@@

Learning the parameters of the \emph{a priori} distribution for each group of
noisy patches amounts to estimating a covariance matrix in a high dimensional space 
and with a limited number of samples. 
%The latter is controlled by the $n$ parameter, which
%should not be too high in order for the Gaussian prior to be valid.
%
The empirical covariance matrix $\widehat C$ converges to the true covariance matrix
$C$ as the number of samples from the underlying Gaussian distribution grows to infinity.
But in our case, the number of samples we can ``draw'' from the Gaussian is limited:
%The Gaussian is only an approximate density model for the set of
%$n$ nearest neighbors of a target patch.
As we increase the number $n$ of neighbors of a patch, we include patches
that are not as similar to the reference patch and the Gaussian model becomes less valid.

Thus there is a trade-off between accuracy of Gaussian model and precision of the
covariance matrix. We will find a compromise to this trade-off empirically  in
\S \ref{sec:parameters}.

The problem of estimating the covariance matrix has received a lot of attention in the
past ten years.
According to \cite{Vershynin2012}, the number of samples needed to achieve a given approximation
error between the empirical and the true covariances grows linearly with the
dimensionality of the space spanned by the data. 
%\pcomment{Comment: Actually, I found results relating the number of samples with
%the dimensionality of the ambient space. 
%However, if the data lies in a $p$-dimensional subspace, one could 
%compute a basis of this subspace from $n > p$ samples, and work in the low-dim
%coordinates to estimate a covariance matrix there. In that way, one could build
%a covariance matrix in the ambient space.}
%\pcomment{Comment: This reference may not be as relevant, in the light of the 
%results known for Gaussian covariance matrices. It would be interesting to see what 
%\cite{bunea2015} says regarding covariance matrices with low-effective rank.}
